(Re)building cooperation: Effects of a cognitive intervention on cooperative behavior in games

### 2023-11-19 article.Rmd
ismail modifs psych science

abstract: |
  Social trust is an important building block of strong social bonds, and its absence is a risk factor for social dysfunction. As such, interventions to foster and strengthen trust-based cooperation are highly desirable. Using the Repeated Trust Game paradigm, we assess the effectiveness of a cognitive intervention derived from Dialectical Behavior Therapy. The intervention's goal was to repair the potential breakdown of cooperation from a pre-programmed, one-off defection by the opponent. Over two games, participants are given the role of the trustee and face what they believe are two different players. In between games, they either receive a brief cognitive intervention or not. Post-intervention, participants showed more cooperative behavior both before and after defection by the opponent. Analysing participants' actions with a hidden Markov model shows that participants in the intervention group had a higher proportion of cooperative states than participants in the control group. This is consistent with participants inferring from the intervention that pro-social and trustworthy behavior may generally provide more beneficial outcomes to them in the long run.

@@

output:
  bookdown::pdf_document2:
   toc: false

---
\small  
^1^ *Department of Experimental Psychology, Division of Psychology and Language Sciences, UCL*.  
^2^ *Division of Psychiatry and Max Planck Centre for Computational Psychiatry and Ageing Research, Queen Square Institute of Neurology, UCL.*

@@

# using some functions dplyr, ggpubr, PairedData and sjPlot. Need to be loaded. 
library(tidyverse)
library(afex)
library(PairedData)
library(multcompView)
library(lsmeans)
library(depmixS4)
library(flextable)
library(gridExtra)
library(forcats)
library(ggsignif)

@@

Determining the other's goals and intentions is fundamental to successful social interaction. Inference of these latent motives is fraught with uncertainty [@barnby2023formalising; @feldmanhall2019resolving], as they are not directly observable and we may require extensive prior experience with the person to infer them. Absent a history of interaction, we may decide to trust that the other's goals and intentions are aligned with our own. Such trust is risky  and if misplaced, it can come at a high cost. But not trusting others is also risky as we may forego mutally beneficial cooperation.

<!-- and any substitution for trust, such as excessive suspicion, over-reliance on self and other maladaptive strategies are likely to add burden and costs to a person. -->

Prior research emphasises the importance of social trust in determining why some people fare better than others physically and mentally [@giordano_trust_2016; @meng_multilevel_2014]. Social trust affects the health status of individuals by reinforcing social support networks, maintaining community norms and facilitating collective action. Research into the determinants of psychopathology has linked trust-based constructs to the emergence of mental health disorders. @fonagy_role_2014 identified epistemic trust -- the belief in the authenticity and personal relevance of interpersonally transmitted knowledge -- as an important function of early attachment relationships. It allows the receiver of social information to let go of their natural self-protective vigilance, which can become pathological hypervigilance after adverse or traumatic experiences and a key factor in the emergence of several mental health disorders [@fonagy_mentalizing_2017]. 
<!-- A well-established paradigm in the study of the emergence and maintenance of trust is the repeated trust game [@joyce_trust_1995]. In this game, one player takes the role of the "investor" and is provided with a fixed endowment at the start of each trial. They get to decide how much of their endowment to invest with the other player taking the role of the "trustee". The amount that is sent is tripled and the trustee gets to decide, in return, how much of the tripled amount to send back to the investor. If the investor sends a non-zero amount, they express trust in the other player for that round, as the amount they will receive back is uncertain. Assuming the trustee sends back more than the initial investment, they signal trustworthiness to the investor and both players make gains. This scenario would constitute a cooperative exchange and would be mutually beneficial. However, cooperation in this setting is prone to be broken through intentional or simply misinterpreted actions that can convey lack of trust or trustworthiness such as low investments and returns [@bendor_when_1991]. In order to repair a damaged cooperative equilibrium, players need to infer that their behavior has violated social norms and offer amends through generous actions at potentially a high cost to them. @king-casas_rupture_2008 links the breakdown of cooperation in trustees with Borderline Personality Disorder to the absence or reduction of activation in brain regions associated with the perception of norm violations.  -->
Since a lack of epistemic trust is a risk factor for social dysfunction [@fonagy_mentalizing_2017], and given the importance of trust for building and maintaining strong social bonds, interventions that foster and strengthen trust-based cooperation would be highly beneficial to society. Such interventions would allow people to more easily repair broken relationships and continue harvesting the benefits of cooperation even in the presence of accidental or intentional social norm violations.

@@

Whilst these results show that it is possible to improve cooperative outcomes at the start of the game, they do not address how to repair a breakdown of trust after intentional or accidental non-cooperative actions by the players. Cooperation in the RTG can easily break down after a transgressive behavior such as a nil or very low investment by the investor or a return by the trustee below the investment sent [@bendor_when_1991]. Such breakdown of cooperation are particularly evident when the trustee suffers from social disorders such as Borderline Personality Disorder [BPD; @lieb_borderline_2004]. Trustees with BPD fail to engage in trust-repairing behaviors such as coaxing the investor by signalling trustworthiness via sending high returns. This failure may be due to BPD trustees not realising that low returns in the game violate social norms [@king-casas_rupture_2008].

To devise interventions to repair social trust, we can take inspiration from cognitive interventions championed by successful therapies that aim to improve social dysfunction in BPD. Psychotherapies such as Mentalisation Based Therapy [@allen_handbook_2006-1] and Dialectical Behavior Therapy [@linehan_cognitive-behavioral_1993] have been shown to improve social skills in BPD [@gunderson_borderline_2018]. However, response to these treatments is highly variable, and determining which interventions are effective for individual patients is challenging [@rudge_mechanisms_2020; @arch_longitudinal_2012]. One promising approach is the study of how specific components of psychotherapeutic treatment affect quantitative markers of behavior such as those inferred through computational models [@huys_computational_2016; @reiter_neuro-cognitive_2021; @dercon2022core]. Combining the use of specific cognitive probes inspired by therapeutic interventions and computational models of behavior may allow us to uncover the cognitive mechanisms targeted by common forms of psychotherapy. In turn, this may provide the basis for choosing effective psychotherapeutic interventions for given individuals.

@@

```{r dataLoad, include=FALSE }
# d_finished <- readRDS("data/d_finished.RDS")
# d_finished$id <- NULL
# saveRDS(d_finished, file="data/d_anonym.RDS")

@@

```{r timeline, fig.cap = "Experiment overview. After playing 15 rounds of the RTG as the trustee, participants were randomized to either be part of the control or intervention condition. They then played the second set of 15 rounds of RTG (again as the trustee) to examine intervention effects, and 7 rounds of a repeated Prisoner’s Dilemma to examine generalization of intervention effects. Finally, participants answered questionnaires and were debriefed.", fig.align='center', out.width="100%", out.height= "120%"}

@@

Participants played a 15-round Repeated Trust Game [@joyce_trust_1995] in the trustee role against a computer-programmed investor. On each round the investor is endowed with 20 units and decides how much of that endowment to invest. This investment is tripled and the trustee then decides how to split this tripled amount between them and the investor. If the trustee returns more than one third of the amount, the investor makes a gain. <!-- The Nash equilibrium for a single-round version is for the investor to send nothing. In the repeated version, rewards for both players are maximised if they build trust and share the benefits of higher investments. An investor who has been rewarded for taking the risk of sending an investment will likely invest more in future rounds. An investor obtaining a low return on their investment may choose to reduce future investment and thereby reduce both players’ gains.-->
```{r, include=FALSE}
Anti_social <-  c(0.1025436430408, 0.1221931236853, 0.1345215238995,
0.1368182252617, 0.1285592471751, 0.1116014322677, 0.0895041220882,
0.0663166721334, 0.0453950277118, 0.0287077419587, 0.0167723588011,
0.0090530028158, 0.0045143317507, 0.0020796744990, 0.0008851094702,
0.0003480141146, 0.0001264134789, 0.0000424213859, 0.0000131513231,
0.0000037665630, 0.0000009965755)

@@

plotinvHMM <- ggplot(response_probs,                            
       aes(x = investment,
           y = probability,
           fill = Investor_state)) +
  geom_bar(stat = "identity",
           position = "dodge") + 
  labs(fill='Latent investor state', x = "Investment", y= "Probability") + 
  theme_bw() + 
  theme(legend.position = "bottom")

@@

  # plotting code...
  ggplot() +
    geom_rect(data = bg_colors, aes(xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, fill = from), alpha = 0.1) +
    geom_line(data = df, aes(x = ret, y = probs, colour = as.factor(to))) +
    facet_wrap(~from, labeller = labeller(from = function(x) paste("From", x, "state on trial t"))) +
    ylim(c(0,1)) +
    scale_fill_manual(values = c("low-trust" = "red", "medium-trust" = "green", "high-trust" = "blue")) +
    labs(x = "Investor's net return on trial t", y = "Transition probability to \nState on trial t+1", color = 'State transitioned to') +
    theme_bw() +
    theme(legend.position = "bottom", 
          axis.title.y = element_text(size = 8, angle = 90, vjust = 1.5))

@@

plotInvTran <- plot_HMM_transitions(3, pars_inv)

@@

```{r HMMPanels, fig.cap="We construct the artificial investor agent by fitting a three-state hidden Markov model to data of human investors engaged in the 10 round Repeated Trust Game against human trustees. From the fitted HMM, we get the distribution of investments by the artificial investor agent conditional on its latent state as shown in Panel A. The fitted HMM also yields the transition probability of the agent to a state on trial t+1 as a function of the net return (difference between the investment sent and the amount received in return) on trial t as shown in Panel B. Each plot in Panel B represents a different starting latent state on trial t, and each line represents the probability of transitioning to a particular state in trial t+1. The policy in each state in Panel A and the way the agent transitions between states as defined in Panel B characterises the artificial agent's behavior in the Repeated Trust Game.", fig.align="center"}

ggpubr::ggarrange(plotinvHMM , plotInvTran, nrow=2, ncol=1, labels = c("A", "B"), common.legend = TRUE, legend = "bottom")

The strategy of the computerised investor was modelled on behavior of human investors in the Repeated Trust Game (RTG)  over 10-rounds with the same (human) opponent. Full detail on the datasets used in the Supplementary Information. Using this data, we estimated a hidden Markov model (HMM) on investors' behavior with three latent states. Each latent state was associated with a state-conditional distribution over the possible investments from 0 to 20 (Figure \ref{fig:HMMPanels}.A). These distributions reflect "low-trust", "medium-trust", or "high-trust". Over rounds, the investor can move between states, and the probability of these transitions was modelled as a function of their net return (i.e return - investment) in the previous round (see Figure \ref{fig:HMMPanels}.B). In order to instigate a potential breakdown of trust, thereby allowing us to probe efforts to repair trust, the computerised agent was programmed to provide a low investment on round 12 (pre-intervention) or round 13 (post-intervention). On all other rounds, the investor's actions were determined by randomly drawing an investment from the state-conditional distribution, with the state over rounds determined by randomly drawing the next state from the state-transition distribution as determined from the net return on the previous round (disregarding the net return immediately after the pre-programmed low investment rounds). The starting state in each instance of the game was the XXX.<!-- MS: what was the starting state? -->

On each round, immediately after being informed of the investment sent, participants in the intervention condition were asked to provide an evaluation of their emotion in terms of valence (from negative to positive) and arousal (from low to high). Participants in the control condition were asked to evaluate the investment in terms of speed (from slow to fast) and magnitude (from low to high). These evaluations were made by clicking on a two-dimensional field with labelled axes.

@@

In the control condition, participants were asked to solve five anagrams ("listen", "triangle", "deductions", "players", "care"). They provided their answers in a free-form text box. The time given to solve the anagrams was the same as that given to respond to questions in the intervention manipulation.

@@

At the start of the experiment (Figure \ref{fig:timeline}), participants provided informed consent and were instructed the study would consist of three phases in which they would face a different other player. Participants were told their goal was to maximise the number of points in all phases. They were not told the number of rounds of each phase. Phase one was a 15 round Repeated Trust Game (RTG) in which participants took the role of trustee, facing the same investor over all 15 rounds. On each round, after being informed about the amount sent by the investor, participants were asked to evaluate their emotion (invervention condition) or the investment (control condition). <!-- provide feedback, with participants in the Intervention condition rating their feeling in terms of valence (from negative to positive) and arousal (from low to high), and participants in the Control condition rating the investment in terms of speed (from slow to fast) and magnitude (from low to high). These ratings were made by indicating a point on a two-dimensional grid.--> Participants then decided on how much of the tripled investment to return to the investor, before continuing to the next round. After completing 15 rounds of the RTG, participants rated how cooperative, selfish, trustworthy and friendly they perceived the investor to be (all on a scale from 1 to 10). After phase one, participants in the intervention condition completed the intervention, and participants in the control condition solved anagrams. Subsequent phase two was similar to phase one, with participants being told they would face a new player. Phase three consisted of 7 rounds of the Repeated Prisoner's Dilemma game (RPD), with participants informed they would face a third player. Participants then completed questionnaires related to mentalising abilities, emotion regulation, and BPD traits (see the supplement for details). They were then asked about the strategy in the games, as well as whether they thought the other players were human or computer agents. They were then debriefed and thanked for their participation.

@@

```{r gamesPlot, echo=FALSE, fig.cap="Averages and standard errors of the trustee's return as a percentage of the multiplied investment received by Condition, Phase, and game round. The blue line shows the returns pre-manipulation and the green line post-manipulation. We note a different reaction to the pre-programmed one-off low investment between the two conditions: Whilst there is a dip in returns pre-manipulation for both conditions,  post manipulation we see higher returns in the intervention condition compared to the dip in returns seen in the control condition in the right panel",fig.align="center", fig.width=6, fig.height = 4}

ggplot(avg_ret_df, aes(x=as.factor(roundNum), y=ret_pct, group=gameNum.f, color = gameNum.f, fill=gameNum.f)) +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "ribbon", aes(ymin=..ymin.., ymax=..ymax..), 
               alpha = 0.3, linetype = 0) +
  labs(x = "Round",
       y = "Percentage Return",
       color = "Game") +
  theme_bw() +
  theme(legend.position = "bottom") +
  facet_wrap(~condition.f) +
  scale_color_manual(values = c("darkblue", "darkgreen")) + scale_fill_manual(values = c("darkblue", "darkgreen"))

@@

```{r, include=FALSE}
pctRet_all <- afex::afex_plot(mod_returns_pct, x = "gameNum.f", trace = "condition.f", dodge = 0.8, error = "within",
            mapping = c("linetype", "shape", "fill"),
            data_geom = geom_violin, 
            data_arg = list(width = 0.5),
            factor_levels = list(gameNum.f = c("Pre", "Post")), legend_title = "Condition") + 
  theme_bw() + 
  labs(y = "Percentage Return", x = "Game") + 
  ggtitle("All rounds percentage returns") + 
  theme(plot.title = element_text(size = 16)) +
  theme(legend.position = "bottom") +
  geom_signif(xmin = 0.8, xmax = 1.8, annotations = "***", y_position = 0.88, tip_length = 0.03, textsize = 3) +  # Intervention Group
  geom_signif(xmin = 1.2, xmax = 2.2, annotations = "*", y_position = 0.80, tip_length = 0.03, textsize = 3)  # Control Group

@@

inv_all <- afex_plot(mod_invs, x = "gameNum.f", trace = "condition.f", dodge = 0.8, error = "within",
            mapping = c("linetype", "shape", "fill"),
            data_geom = geom_violin,
            data_arg = list(width = 0.5),
            factor_levels = list(gameNum.f = c("Pre", "Post")), 
            legend_title = "Condition") + 
  theme_bw() + 
  ggtitle("All rounds investments") +
  labs(y = "HMM Investment", x = "Game") + 
  theme(plot.title = element_text(size = 16)) +
  theme(legend.position = "bottom") +
  geom_signif(xmin = 0.8, xmax = 1.8, annotations = "***", y_position = 1, tip_length = -0.03,textsize = 3) +  # Intervention Group
  geom_signif(xmin = 1.2, xmax = 2.2, annotations = "***", y_position = 18, tip_length = 0.03,textsize = 3)   # Control Group

@@

```{r emoPlotCoax, fig.cap="Self-reported emotion valence and arousal as well as investment z-scores for each round of the Repeated Trust Game averaged across participants in the intervention condition only. The participants' emotion reaction measured during the investor's defection round were similar before (round 12) and after (round 13) the intervention.",fig.align="center", fig.width=7, fig.height = 4}

@@

# # define a response class which only contains the standard slots, no additional slots

>>>>

abstract: |
  Social trust is an important building block of strong social bonds, and its absence is a risk factor for social dysfunction and psychopathology. As such, interventions to foster and strengthen trust-based cooperation are highly desirable. Using the Repeated Trust Game paradigm, we assess the effectiveness of a cognitive intervention derived from Dialectical Behavior Therapy. The intervention's goal was to repair the potential breakdown of cooperation from a pre-programmed, one-off defection by the co-player. Over two games, participants are given the role of the trustee and face what they believe are two different players. In between games, they either receive a brief cognitive intervention or not. Post-intervention, participants showed more cooperative behavior both before and after defection by the co-player. Analysing participants' actions with a hidden Markov model shows that participants in the intervention group had a higher proportion of cooperative states than participants in the control group. This is consistent with participants inferring from the intervention that pro-social and trustworthy behavior may generally provide more beneficial outcomes to them in the long run.

@@

output:
  bookdown::pdf_document2:
    toc: false
header-includes:
  - \usepackage{lscape}
---
\small  
^1^ *Department of Experimental Psychology, Division of Psychology and Language Sciences, UCL*.  
^2^ *Division of Psychiatry and Max Planck Centre for Computational Psychiatry and Ageing Research, Queen Square Institute of Neurology, UCL.*

@@

# using some functions dplyr, ggpubr, PairedData and sjPlot. Need to be loaded. 
library(tidyverse)
library(afex)
library(PairedData)
library(multcompView)
library(lsmeans)
library(depmixS4)
library(flextable)
library(grid)
library(gridExtra)
library(forcats)
library(ggsignif)
library(magick)

@@

Determining the other's goals and intentions is fundamental to successful social interaction. Inference of these latent motives is fraught with uncertainty [@barnby2023formalising; @feldmanhall2019resolving], as they are not directly observable and we may require extensive prior experience with the person to infer them. Absent a history of interaction, we may decide to trust that the other's goals and intentions are aligned with our own. Such trust is risky  and if misplaced, it can come at a high cost. But not trusting others is also risky as we may forego mutually beneficial cooperation.

Prior research emphasises the importance of social trust in determining why some people fare better than others physically and mentally [@giordano_trust_2016; @meng_multilevel_2014]. Social trust affects the health status of individuals by reinforcing social support networks, maintaining community norms and facilitating collective action. Research into the determinants of psychopathology has linked trust-based constructs to the emergence of mental health disorders. @fonagy_role_2014 identified epistemic trust -- the belief in the authenticity and personal relevance of interpersonally transmitted knowledge -- as an important function of early attachment relationships. It allows the receiver of social information to let go of their natural self-protective vigilance, which can become pathological hyper-vigilance after adverse or traumatic experiences and a key factor in the emergence of several mental health disorders [@fonagy_mentalizing_2017; @nolte2023role].

Since a lack of epistemic trust is a risk factor for social dysfunction [@fonagy_mentalizing_2017], and given the importance of trust for building and maintaining strong social bonds, interventions that foster and strengthen trust-based cooperation would be highly beneficial to society. Such interventions would allow people to more easily repair broken relationships and continue harvesting the benefits of cooperation even in the presence of accidental or intentional social norm violations.

@@

Whilst these results show that it is possible to improve cooperative outcomes at the start of the game, they do not address how to repair a breakdown of trust after intentional or accidental non-cooperative actions by the players. Cooperation in the RTG can easily break down after a transgressive behavior such as a nil or very low investment by the investor or a return by the trustee below the investment sent [@bendor_when_1991]. Such breakdown of cooperation are particularly evident when the trustee suffers from social disorders such as Borderline Personality Disorder [BPD; @lieb_borderline_2004]. Trustees with BPD fail to engage in trust-repairing behaviors such as coaxing the investor by signalling trustworthiness via sending high returns. This failure may be due to BPD trustees not realising that low returns in the game violate social norms and the mentalizing difficulties that are associated with appraising others’ intentions [@king-casas_rupture_2008; @hula2018model].

To devise interventions to repair social trust, we can take inspiration from cognitive interventions championed by successful therapies that aim to improve social dysfunction in BPD. Psychotherapies such as Mentalisation Based Therapy [@allen_handbook_2006-1] and Dialectical Behavior Therapy [@linehan_cognitive-behavioral_1993] have been shown to improve social skills in BPD [@gunderson_borderline_2018]. However, response to these treatments is highly variable, and determining which interventions are effective for particular sub-groups of patients is challenging [@rudge_mechanisms_2020; @arch_longitudinal_2012]. One promising approach is the study of how specific components of psychotherapeutic treatment affect quantitative markers of behavior such as those inferred through computational models [@huys_computational_2016; @reiter_neuro-cognitive_2021; @dercon2022core]. Combining the use of specific cognitive probes inspired by therapeutic interventions and computational models of behavior may allow us to uncover the cognitive mechanisms targeted by common forms of psychotherapy. In turn, this may provide the basis for choosing effective psychotherapeutic interventions for given individuals.

@@

```{r dataLoad, include=FALSE }

@@

```{r timeline, include=F, fig.cap = "Experiment overview. After playing 15 rounds of the RTG as the trustee, participants were randomized to either be part of the control or intervention condition. They then played the second set of 15 rounds of RTG (again as the trustee) to examine intervention effects, and 7 rounds of a repeated Prisoner’s Dilemma to examine generalization of intervention effects. Finally, participants answered questionnaires and were debriefed.", fig.align='center', out.width="100%", out.height= "120%"}

@@

Participants played a 15-round Repeated Trust Game [@joyce_trust_1995] in the trustee role against a computer-programmed investor. On each round the investor is endowed with 20 units and decides how much of that endowment to invest. This investment is tripled and the trustee then decides how to split this tripled amount between them and the investor. If the trustee returns more than one third of the amount, the investor makes a gain. Each player was represented with an icon with the participant always on the left of the screen and the co-player on the right (Figure \ref{fig:HMMPanels}.A). The participants were able to choose the icon that represents them at the start of the experiment. The icon representing the co-player changed at the start of each new game, to simulate a new interaction partner. Participants were not told they were facing computerised co-players.

The strategy of the computerised investor was modelled on behavior of human investors in the Repeated Trust Game (RTG)  over 10-rounds with the same (human) co-player. Full detail on the datasets used in the Supplementary Information. Using this data, we estimated a hidden Markov model (HMM) on investors' behavior with three latent states. Each latent state was associated with a state-conditional distribution over the possible investments from 0 to 20 (Figure \ref{fig:HMMPanels}.C). These distributions reflect "low-trust", "medium-trust", or "high-trust". Over rounds, the investor can move between states, and the probability of these transitions was modelled as a function of their net return (i.e return - investment) in the previous round (see Figure \ref{fig:HMMPanels}.D). In order to instigate a potential breakdown of trust, thereby allowing us to probe efforts to repair trust, the computerised agent was programmed to provide a low investment on round 12 (pre-intervention) or round 13 (post-intervention). On all other rounds, the investor's actions were determined by randomly drawing an investment from the state-conditional distribution, with the state over rounds determined by randomly drawing the next state from the state-transition distribution as determined from the net return on the previous round (disregarding the net return immediately after the pre-programmed low investment rounds). The initial state for the HMM investor in each instance of the game was the "mid-trust" state.

On each round, immediately after being informed of the investment sent, participants in the intervention condition were asked to provide an evaluation of their emotion in terms of valence (from negative to positive) and arousal (from low to high). Participants in the control condition were asked to evaluate the investment in terms of speed (from slow to fast) and magnitude (from low to high). These evaluations were made by clicking on a two-dimensional field with labelled axes as shown in Figure \ref{fig:HMMPanels}.C for the intervention condition.

```{r, include=FALSE}
Anti_social <-  c(0.1025436430408, 0.1221931236853, 0.1345215238995,
0.1368182252617, 0.1285592471751, 0.1116014322677, 0.0895041220882,
0.0663166721334, 0.0453950277118, 0.0287077419587, 0.0167723588011,
0.0090530028158, 0.0045143317507, 0.0020796744990, 0.0008851094702,
0.0003480141146, 0.0001264134789, 0.0000424213859, 0.0000131513231,
0.0000037665630, 0.0000009965755)

@@

plotinvHMM <- ggplot(response_probs,                            
       aes(x = investment,
           y = probability,
           fill = Investor_state)) +
  geom_bar(stat = "identity",
           position = "dodge") + 
  labs(fill='Latent investor state', x = "Investment", y= "Probability") + 
  theme_bw() + 
  theme(legend.position = "bottom",  legend.text = element_text(size = 12),  legend.title = element_text(size = 14))

plotinvHMM
```

@@

  # plotting code...
  ggplot() +
    geom_rect(data = bg_colors, aes(xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, fill = from), alpha = 0.1) +
    geom_line(data = df, aes(x = ret, y = probs, colour = as.factor(to))) +
    facet_wrap(~from, labeller = labeller(from = function(x) paste("From", x, "state on trial t"))) +
    ylim(c(0,1)) +
    scale_fill_manual(values = c("low-trust" = "red", "medium-trust" = "green", "high-trust" = "blue"),
                    name = "From state") +  # Changed legend title for 'fill' here
    scale_color_manual(values = c("low-trust" = "red", "medium-trust" = "green", "high-trust" = "blue"),
                       labels = c("low-trust", "medium-trust", "high-trust"),
                       name = "State transitioned to") +
    labs(x = "Investor's net return on trial t", y = "Transition probability to \nState on trial t+1", color = 'State transitioned to') +
    theme_bw() +
    theme(legend.position = "bottom",
          legend.text = element_text(size = 12),
          legend.key.size = unit(1, 'lines'),
          legend.spacing.x = unit(0.1, 'in'),
          legend.title = element_text(size = 14),
          legend.margin = margin(t = 0.2, b = 0, unit = 'cm'),
          plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "cm"),
          strip.text = element_text(size = 10),
          legend.box = "vertical" # Arrange legends vertically
    ) +
    guides(fill = guide_legend(order = 1, title.position = "left", title.hjust = 0.3),
           color = guide_legend(order = 2, title.position = "left", title.hjust = 0.3))
}

@@

plotInvTran <- plot_HMM_transitions(3, pars_inv) 
print(plotInvTran)

@@

\begin{landscape}
```{r HMMPanels, fig.cap= "\\small{A: Screenshot of the repeated Trust Game. The game is played in dyads, with one player assigned the role of the investor and the other player that of the trustee. The investor is endowed with 20 units at the start of each round. They need to decide how much of that endowment they want to invest with the trustee. The investment is then multiplied by a factor of 3 and sent to the trustee who needs to decide how much of the multiplied investment they want to send back to the investor. Shown here is the stage at which the trustee makes a decision of how much to send back the investor. Immediately before making this decision, participant are shown a two dimensional grid (Panel C) where they need to report their reaction to the investment. Those in the intervention condition were asked questions about their emotional state and those in the control condition were asked to report features of the investment unrelated to their emotion. Panels B - D: We construct the artificial investor agent by fitting a three-state hidden Markov model to data of human investors engaged in the 10 round Repeated Trust Game against human trustees. From the fitted HMM, we get the distribution of investments by the artificial investor agent conditional on its latent state as shown in Panel C. The fitted HMM also yields the transition probability of the agent to a state on trial t+1 as a function of the net return (difference between the investment sent and the amount received in return) on trial t as shown in Panel D. Each plot in Panel D represents a different starting latent state on trial t, and each line represents the probability of transitioning to a particular state in trial t+1.}",echo=FALSE, message=FALSE, warning=FALSE, out.width='1\\linewidth', out.height='1\\textheight', fig.align='center'}

library(magick)

# Read in the images
image1 <- image_read("figures/trust_game.png")
image2 <- image_read("figures/HMMinvPolicy2.png")
image3 <- image_read("figures/emotion_grid.png")
image4 <- image_read("figures/HMMinvTrans2.png")

# Determine the target width for all images, for example, the width of the widest image
target_width <- max(image_info(image1)$width, image_info(image2)$width, 
                    image_info(image3)$width, image_info(image4)$width)

target_height <- max(image_info(image1)$height, image_info(image2)$height, 
                    image_info(image3)$height, image_info(image4)$height)

# Scale images to have the same width
image1 <- image_scale(image1, paste0(target_width, "x", target_height))
image2 <- image_scale(image2, paste0(target_width, "x", target_height))
image3 <- image_scale(image3, paste0(target_width, "x", target_height))
image4 <- image_scale(image4, paste0(target_width, "x", target_height))

# Annotate each image with a label
image1 <- image_annotate(image1, text = "A", location = "+200+10", size = 90, color = "black", gravity = "northwest")
image2 <- image_annotate(image2, text = "B", location = "+0+10", size = 90, color = "black", gravity = "northwest")
image3 <- image_annotate(image3, text = "C", location = "-0+10", size = 90, color = "black", gravity = "northwest")
image4 <- image_annotate(image4, text = "D", location = "+0+10", size = 90, color = "black", gravity = "northwest")

# Assuming a 300 DPI resolution for good print quality
dpi <- 300

# Spacing of 0.1 inches converted to pixels
spacing <- 0.05 * dpi

# Calculating the width and height for each image
width_per_image <- ((12 - 0.1) / 2) * dpi
height_per_image <- ((7.4 - 0.1) / 2) * dpi
# The geometry string for each image, including spacing (assuming no borders)
geometry_string <- paste0(width_per_image, 'x', height_per_image, '+', spacing, '+', spacing)

# Now create the montage with the specified geometry
combined_image <- image_montage(
  c(image1, image2, image3, image4), 
  tile = "2x2",
  geometry = geometry_string
)

# Display the combined image
plot(combined_image)
```
\end{landscape}

@@

In order to separate any general practice effects from the effect of the intervention, we added a control condition in which participants  were asked to solve five anagrams ("listen", "triangle", "deductions", "players", "care"). They provided their answers in a free-form text box. The time given to solve the anagrams was the same as that given to respond to questions in the intervention manipulation.

@@

At the start of the experiment, participants provided informed consent and were instructed the study would consist of three phases in which they would face a different other player. Participants were told their goal was to maximise the number of points in all phases. They were not told the number of rounds of each phase. Phase one was a 15 round Repeated Trust Game (RTG) in which participants took the role of trustee, facing the same investor over all 15 rounds. On each round, after being informed about the amount sent by the investor, participants were asked to evaluate their emotion (intervention condition) or the investment (control condition). <!-- provide feedback, with participants in the Intervention condition rating their feeling in terms of valence (from negative to positive) and arousal (from low to high), and participants in the Control condition rating the investment in terms of speed (from slow to fast) and magnitude (from low to high). These ratings were made by indicating a point on a two-dimensional grid.--> Participants then decided on how much of the tripled investment to return to the investor, before continuing to the next round. After completing 15 rounds of the RTG, participants rated how cooperative, selfish, trustworthy and friendly they perceived the investor to be (all on a scale from 1 to 10). After phase one, participants in the intervention condition completed the intervention, and participants in the control condition solved anagrams. Subsequent phase two was similar to phase one, with participants being told they would face a new player. Phase three consisted of 7 rounds of the Repeated Prisoner's Dilemma game (RPD), with participants informed they would face a third player. Participants then completed questionnaires related to mentalising abilities, emotion regulation, and BPD traits (see the supplement for details). They were then asked about the strategy in the games, as well as whether they thought the other players were human or computer agents. They were then debriefed and thanked for their participation.

@@

```{r gamesPlot, include=FALSE, echo=FALSE, fig.cap="Averages and standard errors of the trustee's return as a percentage of the multiplied investment received by Condition, Phase, and game round. The blue line shows the returns pre-manipulation and the green line post-manipulation. We note a different reaction to the pre-programmed one-off low investment between the two conditions: Whilst there is a dip in returns pre-manipulation for both conditions,  post manipulation we see higher returns in the intervention condition compared to the dip in returns seen in the control condition in the right panel",fig.align="center", fig.width=6, fig.height = 4}

# ggplot(avg_ret_df, aes(x=as.factor(roundNum), y=ret_pct, group=gameNum.f, color = gameNum.f, fill=gameNum.f)) +
#   stat_summary(fun = mean, geom = "line") +
#   stat_summary(fun.data = mean_se, geom = "ribbon", aes(ymin=..ymin.., ymax=..ymax..), 
#                alpha = 0.3, linetype = 0) +
#   labs(x = "Round",
#        y = "Percentage Return",
#        color = "Game") +
#   theme_bw() +
#   theme(legend.position = "bottom") +
#   facet_wrap(~condition.f) +
#   scale_color_manual(values = c("darkblue", "darkgreen")) + scale_fill_manual(values = c("darkblue", "darkgreen"))

ggplot(avg_ret_df, aes(x=as.factor(roundNum), y=ret_pct, group=gameNum.f, color = gameNum.f, fill=gameNum.f)) +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "ribbon", aes(ymin=..ymin.., ymax=..ymax..),
               alpha = 0.3, linetype = 0) +
  labs(x = "Round",
       y = "Percentage Return",
       color = "Trust Game Phase") +
  theme_bw() +
  theme(legend.position = "bottom",
        legend.key = element_blank(),
        #legend.key = element_rect(fill = NA), # Make legend key background transparent
        legend.margin = margin(t = 0, b = 0, unit = "pt"), # Adjust top and bottom margin of each legend
        legend.box.margin = margin(t = 0, b = 0, unit = "pt")) + # Adjust top and bottom margin of the legend box
  facet_wrap(~condition.f) +
  scale_color_manual(values = c("darkblue", "darkgreen"),
                     labels = c("Pre-manipulation", "Post-manipulation")) +
  scale_fill_manual(values = c("darkblue", "darkgreen"),
                    guide = "none") + # Hide the fill legend
  guides(color = guide_legend(override.aes = list(fill = NA))) # Remove fill for color legend keys

@@

```{r, include=FALSE}
pctRet_all <- afex::afex_plot(mod_returns_pct, x = "gameNum.f", trace = "condition.f", dodge = 0.8, error = "within",
            mapping = c("linetype", "shape", "fill"),
            data_geom = geom_violin, 
            data_arg = list(width = 0.5),
            factor_levels = list(gameNum.f = c("Pre-manipulation", "Post-manipulation")), legend_title = "Condition") + 
  theme_bw() + 
  labs(y = "Percentage Return", x = "Trust Game Phase") + 
  ggtitle("All rounds percentage returns") + 
  theme(plot.title = element_text(size = 16)) +
  theme(legend.position = "bottom") +
  geom_signif(xmin = 0.8, xmax = 1.8, annotations = "***", y_position = 0.88, tip_length = 0.03, textsize = 3) +  # Intervention Group
  geom_signif(xmin = 1.2, xmax = 2.2, annotations = "*", y_position = 0.80, tip_length = 0.03, textsize = 3)  # Control Group

@@

inv_all <- afex_plot(mod_invs, x = "gameNum.f", trace = "condition.f", dodge = 0.8, error = "within",
            mapping = c("linetype", "shape", "fill"),
            data_geom = geom_violin,
            data_arg = list(width = 0.5),
            factor_levels = list(gameNum.f = c("Pre-manipulation", "Post-manipulation")), 
            legend_title = "Condition") + 
  theme_bw() + 
  ggtitle("All rounds investments") +
  labs(y = "HMM Investment", x = "Trust Game Phase") + 
  theme(plot.title = element_text(size = 16)) +
  theme(legend.position = "bottom") +
  geom_signif(xmin = 0.8, xmax = 1.8, annotations = "***", y_position = 1, tip_length = -0.03,textsize = 3) +  # Intervention Group
  geom_signif(xmin = 1.2, xmax = 2.2, annotations = "***", y_position = 18, tip_length = 0.03,textsize = 3)   # Control Group

@@

```{r emoPlotCoax, include=F, fig.cap="Self-reported emotion valence and arousal as well as investment z-scores for each round of the Repeated Trust Game averaged across participants in the intervention condition only. The participants' emotion reaction measured during the investor's defection round were similar before (round 12) and after (round 13) the intervention.",fig.align="center", fig.width=7, fig.height = 4}

@@
